---
title: "Majorproject"
author: "Caitlin Sutherland"
date: "09/04/2020"
output: html_document
---

Load the dataset. 

```{r}
SIAdata=read.csv("SIAData_edit.csv")
summary(SIAdata)
install.packages("tidyverse")
library(tidyverse)
library(vegan)
```

The response variables are: SpFPOM, SpCPOM, SpCG, FPOM, CPOM, and CG. 

##PART I: CHECKING FOR NORMALITY

Checking to see if data is normally distributed. Will look at histograms and qqplots to explore the data visually, then run Shapiro-Wilk normality test to confirm any suspicions we may have. 

```{r}
hist(SIAdata$SpFPOM)
hist(SIAdata$SpCPOM)
hist(SIAdata$SpCG)
hist(SIAdata$FPOM)
hist(SIAdata$CPOM)
hist(SIAdata$CG)
```

Based on the histograms, the most normally distributed variable seems to be SpCG, followed by CG. 

```{r}
qqnorm(SIAdata$SpFPOM)
qqline(SIAdata$SpFPOM,col="red",lwd=2)

qqnorm(SIAdata$SpCPOM)
qqline(SIAdata$SpCPOM,col="red",lwd=2)

qqnorm(SIAdata$SpCG)
qqline(SIAdata$SpCG,col="red",lwd=2)

qqnorm(SIAdata$FPOM)
qqline(SIAdata$FPOM,col="red",lwd=2)

qqnorm(SIAdata$CPOM)
qqline(SIAdata$CPOM,col="red",lwd=2)

qqnorm(SIAdata$CG)
qqline(SIAdata$CG,col="red",lwd=2)
```

Based on the QQplots, the only variable with normal distribution is SpCG. 

Will also perform Shapiro-Wilk normality test. The null hypothesis is that the data is normally distributed; if the S-W test p < 0.05, we will reject the null hypothesis and determine that the data is NOT normally distributed. 

```{r}
shapiro.test(SIAdata$SpFPOM)
shapiro.test(SIAdata$SpCPOM)
shapiro.test(SIAdata$SpCG)
shapiro.test(SIAdata$FPOM)
shapiro.test(SIAdata$CPOM)
shapiro.test(SIAdata$CG)
```

The Shapiro-Wilk normality test confirms that the only normal dataset is that of the response variable SpCG.  

The data will need to be transformed before parametric tests are performed. 

```{r}
tf.SPFPOM=log2(SIAdata$SpFPOM)
tf.SPCPOM=log2(SIAdata$SpCPOM)
tf.SPCG=log2(SIAdata$SpCG)
tf.FPOM=log2(SIAdata$FPOM)
tf.CPOM=log2(SIAdata$CPOM)
tf.CG=log2(SIAdata$CG)

hist(tf.SPFPOM)
hist(tf.SPCPOM)
hist(tf.SPCG)
hist(tf.FPOM)
hist(tf.CPOM)
hist(tf.CG)

shapiro.test(tf.SPFPOM)
shapiro.test(tf.SPCPOM)
shapiro.test(tf.SPCG)
shapiro.test(tf.FPOM)
shapiro.test(tf.CPOM)
shapiro.test(tf.CG)
```


All response variables with non-normal distribution have been transformed. Will now compile transformed vectors into one dataframe. 

```{r}
tf.data=data.frame('SiteID'=SIAdata$SiteID,'CropCover'=SIAdata$CropCover,'LivestockDens'=SIAdata$LivestockDens,'WWDens'=SIAdata$WWDens,tf.SPCPOM,tf.SPFPOM,tf.SPCG,tf.FPOM,tf.CPOM,tf.CG)

```

##PART II: Correlation Analysis 

Run the correlation analyses. 

Correlations assume a linear relationship. Should take a look at scatterplots and see how data looks. 

```{r}
plot(tf.data$CropCover,tf.data$tf.SPCPOM)
plot(tf.data$CropCover,tf.data$tf.SPFPOM)
plot(tf.data$CropCover,tf.data$tf.SPCG)
plot(tf.data$CropCover,tf.data$tf.CPOM)
plot(tf.data$CropCover,tf.data$tf.FPOM)
plot(tf.data$CropCover,tf.data$tf.CG)
plot(tf.data$LivestockDens,tf.data$tf.SPCPOM)
plot(tf.data$LivestockDens,tf.data$tf.SPFPOM)
plot(tf.data$LivestockDens,tf.data$tf.SPCG)
plot(tf.data$LivestockDens,tf.data$tf.CPOM)
plot(tf.data$LivestockDens,tf.data$tf.FPOM)
plot(tf.data$LivestockDens,tf.data$tf.CG)
plot(tf.data$WWDens,tf.data$tf.SPCPOM)
plot(tf.data$WWDens,tf.data$tf.SPFPOM)
plot(tf.data$WWDens,tf.data$tf.SPCG)
plot(tf.data$WWDens,tf.data$tf.CPOM)
plot(tf.data$WWDens,tf.data$tf.FPOM)
plot(tf.data$WWDens,tf.data$tf.CG)
```

i) Crop cover vs δ15N in coarse particulate organic matter, fine particulate organic matter and collector-gatherer benthic macroinvertebrates

```{r}
cor(tf.data$CropCover,tf.data$tf.SPCPOM,use="complete.obs",method="pearson")
cor(tf.data$CropCover,tf.data$tf.SPFPOM,use="complete.obs",method="pearson")
cor(tf.data$CropCover,tf.data$tf.SPCG,use="complete.obs",method="pearson")
cor(tf.data$CropCover,tf.data$tf.CPOM,use="complete.obs",method="pearson")
cor(tf.data$CropCover,tf.data$tf.FPOM,use="complete.obs",method="pearson")
cor(tf.data$CropCover,tf.data$tf.CG,use="complete.obs",method="pearson")

```

i) Livestock density vs δ15N in coarse particulate organic matter, fine particulate organic matter and collector-gatherer benthic macroinvertebrates

```{r}
cor(tf.data$LivestockDens,tf.data$tf.SPCPOM,use="complete.obs",method="pearson")
cor(tf.data$LivestockDens,tf.data$tf.SPFPOM,use="complete.obs",method="pearson")
cor(tf.data$LivestockDens,tf.data$tf.SPCG,use="complete.obs",method="pearson")
cor(tf.data$LivestockDens,tf.data$tf.CPOM,use="complete.obs",method="pearson")
cor(tf.data$LivestockDens,tf.data$tf.FPOM,use="complete.obs",method="pearson")
cor(tf.data$LivestockDens,tf.data$tf.CG,use="complete.obs",method="pearson")
```

iii) Wastewater release density vs δ15N in coarse particulate organic matter, fine particulate organic matter and collector-gatherer benthic macroinvertebrates

```{r}
cor(tf.data$WWDens,tf.data$tf.SPCPOM,use="complete.obs",method="pearson")
cor(tf.data$WWDens,tf.data$tf.SPFPOM,use="complete.obs",method="pearson")
cor(tf.data$WWDens,tf.data$tf.SPCG,use="complete.obs",method="pearson")
cor(tf.data$WWDens,tf.data$tf.CPOM,use="complete.obs",method="pearson")
cor(tf.data$WWDens,tf.data$tf.FPOM,use="complete.obs",method="pearson")
cor(tf.data$WWDens,tf.data$tf.CG,use="complete.obs",method="pearson")
```

Need to perform a hypothesis test on the correlation coefficients.
Null hypothesis: there is no linear relationship between the two variables; ρ = 0.
Alternative hypothesis: there is a linear relationship between the two variables; ρ ≠ 0.

Go back and run glm function - could have used this to extract correlation coefficient by taking square root of r2 value. Is beneficial to use because it returns a p-value. We can look a this p value of the slope of the regression parameter and determine the significance of the correlation result. 

```{r}
lm1=glm(tf.data$CropCover~tf.data$tf.SPCPOM)
summary(lm1)
lm2=glm(tf.data$CropCover~tf.data$tf.SPFPOM)
summary(lm2)
lm3=glm(tf.data$CropCover~tf.data$tf.SPCG)
summary(lm3)
lm4=glm(tf.data$CropCover~tf.data$tf.CPOM)
summary(lm4)
lm5=glm(tf.data$CropCover~tf.data$tf.FPOM)
summary(lm5)
lm6=glm(tf.data$CropCover~tf.data$tf.CG)
summary(lm6)
```

```{r}
lm7=glm(tf.data$LivestockDens~tf.data$tf.SPCPOM)
summary(lm7)
lm8=glm(tf.data$LivestockDens~tf.data$tf.SPFPOM)
summary(lm8)
lm9=glm(tf.data$LivestockDens~tf.data$tf.SPCG)
summary(lm9)
lm10=glm(tf.data$LivestockDens~tf.data$tf.CPOM)
summary(lm10)
lm11=glm(tf.data$LivestockDens~tf.data$tf.FPOM)
summary(lm11)
lm12=glm(tf.data$LivestockDens~tf.data$tf.CG)
summary(lm12)
```

```{r}
lm13=glm(tf.data$WWDens~tf.data$tf.SPCPOM)
summary(lm13)
lm14=glm(tf.data$WWDens~tf.data$tf.SPFPOM)
summary(lm14)
lm15=glm(tf.data$WWDens~tf.data$tf.SPCG)
summary(lm15)
lm16=glm(tf.data$WWDens~tf.data$tf.CPOM)
summary(lm16)
lm17=glm(tf.data$WWDens~tf.data$tf.FPOM)
summary(lm17)
lm18=glm(tf.data$WWDens~tf.data$tf.CG)
summary(lm18)
```

##PART III: Cluster analysis

Performing a cluster analysis. Need to delete missing data and standardize the variables. 

```{r}
library(ade4)
library(vegan)
library(gclus)
library(ape)

clean.tf.data=na.omit(tf.data) #listwise deletion of missing data
rownames(clean.tf.data)=clean.tf.data[,1]
clean.tf.data=clean.tf.data[,-1]
cluster=hclust(dist(scale(clean.tf.data)), "ward.D")
plot(cluster)
```
